---
title: "Homework 3"
author: "Kelly Odion, Anna Calle"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Part 0. load packages and csv
```{r}
library(tidyverse)

votes <- read_csv("HW3Data.csv")

```


Part 1. Linear probability model 
1. Create a linear probability model that predicts a respondent’s probability of voting ‘yes’ on the
ballot based on their age, income, NEP score, the risk reduction offered by the program, and the
cost of the program to that respondent. Show the model and interpret the regression
coefficients.
```{r}
probability_yes <- glm(vote ~ age + income + NEP + bid + risk, family = "binomial", data = votes)

summary(probability_yes) 

```

2. Based on this regression, what can you say about the value of a single prevented whale death?
(Hint: Think about how risk reduction for all whales translates into the number of whale deaths
avoided)

For 100% increase in risk reduction, people are .4% more likely to vote yes

Risk reduction = 100% 

Cost of implementation: $7 million

3. Pick three arbitrary respondents. For each of those respondents, estimate their WTP for a VSR
program offering 60% risk reduction.

Respondent 1: 
Risk = 20, Bid = 80, Vote = 0, NEP = 15; 

Repondent 2: 
Risk = 40, Bid = 100, vote = 1, NEP = 15; 
 
Respondent 3: 
Risk = 80, Bid = 40, vote = 1, NEP = 32;


4. Now repeat this process and estimate the mean WTP for a VSR program offering 60% risk
reduction among Santa Barbara County households.
5. If there are 150,000 households in Santa Barbara County, estimate the total benefit of a VSR
program to these households.
6. Based only on these benefits, do the benefits of the VSR program outweigh the costs?